---
title: "model help"
author: "David S, Mason"
date: "3/12/2022"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
##### SEED DISPERSAL PROJECT #####

# H01: as food diversity increases, bird and seed richness increase
# H02: as food diversity increases, bird and seed observations increase


##### PACKAGES #####
####################
library(tidyverse)
library(lme4)
library(lubridate)
library(lmerTest)
library(emmeans)
library(car)
library(MuMIn)
library(agridat)
library(DHARMa)
library(glmmTMB)
library(fitdistrplus)


###### DATA ########
####################
bird.dat.og <- read.csv(file = "data/bird_data2.0.csv",
                   header = TRUE)
head(bird.dat.og)

seed.dat.og <- read.csv(file = "data/seed_traps2.0.csv",
                   header = TRUE)
head(seed.dat.og)


### ACCLIM PER #####
####################

# Birds
bird.dat <- dplyr::select(bird.dat.og, -c("camera.card", "X", "notes", "photo.number"))
bird.dat$date <- as.Date(with(bird.dat, paste(year, month, day, sep = "-")), "%Y-%m-%d")

bird.week <- bird.dat %>% 
	group_by(week = week(date)) %>% 
	summarise(value = sum(presence))

bird.week
# looks like a 2 week acclimitizaton period
# observatons really picked up in January (observations are
# essentially increasing with time throughout our sampling period)

bird.dat.drop <- bird.dat %>% 
	mutate(week = week(date)) %>% 
	filter(week != 48 & week != 47)

# remove NA's
bird.clean <- bird.dat.drop %>% 
		 drop_na(species)

bird.clean$treatment <- as.factor(bird.clean$treatment)

# Seeds
seed.dat <- pivot_longer(seed.dat.og, cols = 5:35, names_to = "SPECIES",
												 values_to = "SEEDS")

seed.dat$DATE <- as.Date(seed.dat$DATE)

seed.dat <- seed.dat %>% 
	mutate(week = week(DATE)) %>% 
	filter(week != 48 & week != 47)


## SUMMARIZE DATA ##
####################
#### SUMMARIZE #####
####################
# Birds
# count of individuals by species at each treatment and site
bird.sps.count.by.treat <- bird.clean %>% 
  group_by(site, treatment, species) %>% 
  summarise(count = n())
 
# total count of individuals by treatment and site
bird.obs <- bird.clean %>% 
  group_by(site, treatment) %>% 
  summarise(count = n())

bird.obs <- as.data.frame(bird.obs) %>% 
	dplyr::select(treatment,site,count)

missing <- data.frame(site = "six", 
											treatment = "4",
											count = 0)

bird.obs <- rbind(bird.obs,missing)

sites <- c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4,5,5,5,5,6,6,6,6,7,7,7,7,
					 8,8,8,8,9,9,9,9,10,10,10,10)
treatment <- c(0,4,8,12,0,4,8,12,0,4,8,12,0,4,8,12,
							 0,4,8,12,0,4,8,12,0,4,8,12,0,4,8,12,
							 0,4,8,12,0,4,8,12)

bird.rich <- as.data.frame(cbind(sites,treatment))
bird.rich$richness <- c(2,2,3,3,2,2,2,3,3,3,4,4,3,3,3,3,4,4,5,5,
									 2,0,2,1,1,2,2,1,4,4,4,4,2,2,2,2,1,1,1,1)
bird.rich$treatment <- as.factor(bird.rich$treatment)
bird.rich <- as.data.frame(bird.rich) %>% 
	dplyr::select(treatment,sites,richness)

# Seeds
seed.obs <- seed.dat %>% 
  dplyr::group_by(TREATMENT, BLOCK) %>% 
  dplyr::summarise(sum(SEEDS)) 
seed.obs <- dplyr::rename(seed.obs, SEEDS = "sum(SEEDS)")

# total count of species by treatment and site
seed.species.count <- seed.dat %>% 
  group_by(TREATMENT, BLOCK, SPECIES) %>% 
  summarise(sum(SEEDS)) 

seed.species.count <- dplyr::rename(seed.species.count, SEEDS = "sum(SEEDS)")

seed.species.count$TREATMENT <- factor(seed.species.count$TREATMENT, 
																		levels=c('Control', 'Low', 'Medium', 'High'))

seed.species.count <- seed.species.count %>% 
	filter(SEEDS > 0)

seed.rich <- as.data.frame(cbind(sites,treatment))
seed.rich$richness <- c(1,2,5,2,0,0,0,2,1,0,0,5,0,3,4,3,0,2,2,3,
												1,1,2,2,1,1,1,1,0,1,1,3,0,0,4,4,2,0,0,0)
seed.rich <- as.data.frame(seed.rich) %>% 
	dplyr::select(treatment,sites,richness)

####################
```

# Seed detections distribution
```{r, echo=FALSE}
plotdist(seed.obs$SEEDS, histo = TRUE, demp = TRUE)
descdist(seed.obs$SEEDS, discrete=TRUE, boot=500) # poisson
```

# Negative binomial and poisson model fits for seed detections
```{r}
# neg binom fit
seeds.obs.fit.nb <- glmer.nb(SEEDS ~ TREATMENT + (1 | BLOCK),
                             data = seed.obs)
# poisson fit
seed.obs.fit.pois <- glmer(SEEDS ~ TREATMENT + (1 | BLOCK),
                            data = seed.obs, family = poisson(link = "log"))
```

# Check some assumptions (seed detection models)
```{r, echo=FALSE}
plot(SEEDS ~ as.numeric(TREATMENT), 
     data = seed.obs, main = "Seed detections by treatment")
plot(residuals(seeds.obs.fit.nb) ~ as.numeric(seed.obs$TREATMENT),
     main = "Negative binomial residuals by treatment")
abline(a = 0, b = 0, col = "blue", lwd = 2)
plot(residuals(seed.obs.fit.pois) ~ as.numeric(seed.obs$TREATMENT), main = "Poisson residuals by treatment")
abline(a = 0, b = 0, col = "blue", lwd = 2)
hist(residuals(seeds.obs.fit.nb), main = "Negative binomial res")
hist(residuals(seed.obs.fit.pois), main = "Poisson res")
```

# Plot residuals from prediction to test for assumptions (seed detection models)
```{r, echo=FALSE}
resfit.nb <- resid(seeds.obs.fit.nb)
hist(resfit.nb, main = "Negative binomial")
plot(seed.obs$SEEDS, resfit.nb, 
     ylab = "Residuals", 
     xlab = "Count",
     main = "Negative binomial") 
abline(0, 0)      

resfit.pois <- resid(seed.obs.fit.pois)
hist(resfit.pois, main = "Poisson")
plot(seed.obs$SEEDS, resfit.pois, 
     ylab = "Residuals", 
     xlab = "Count",
     main = "Poisson") 
abline(0, 0)      
```

# Seed richness distribution
```{r, echo=FALSE}
plotdist(seed.rich$richness, histo = TRUE, demp = TRUE)
descdist(seed.rich$richness, discrete=TRUE, boot=500) # poisson
```

# Negative binomial and poisson model fits for seed richness models
```{r}
# neg binom fit
seeds.rich.fit.nb <- glmer.nb(richness ~ treatment + (1 | sites),
                             data = seed.rich) # iteration limit reached
# poisson fit
seeds.rich.fit.pois <- glmer(richness ~ treatment + (1 | sites),
                            data = seed.rich, family = poisson(link = "log"))
```

# Check some assumptions (seed richness models)
```{r, echo=FALSE}
plot(richness ~ as.numeric(treatment), 
     data = seed.rich, main = "Seed richness by treatment")
plot(residuals(seeds.rich.fit.nb) ~ as.numeric(seed.rich$treatment), main = "Negative binomial residuals by treatment")
abline(a = 0, b = 0, col = "blue", lwd = 2)
plot(residuals(seeds.rich.fit.pois) ~ as.numeric(seed.rich$treatment), main = "Poisson residuals by treatment")
abline(a = 0, b = 0, col = "blue", lwd = 2)
hist(residuals(seeds.rich.fit.nb), main = "Negative binomial res")
hist(residuals(seeds.rich.fit.pois), main = "Poisson res")
```

# Plot residuals from prediction to test for assumptions (seed richness models)
```{r, echo=FALSE}
resfit.nb <- resid(seeds.rich.fit.nb)
hist(resfit.nb, main = "Negative binomial")
plot(seed.rich$richness, resfit.nb, 
     ylab = "Residuals", 
     xlab = "Count",
     main = "Negative binomial") 
abline(0, 0)      

resfit.pois <- resid(seeds.rich.fit.pois)
hist(resfit.pois, main = "Poisson")
plot(seed.rich$richness, resfit.pois, 
     ylab = "Residuals", 
     xlab = "Count",
     main = "Poisson") 
abline(0, 0)      
```
